Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-07-10 Paper/paper.tex
Update
Changes from Gabor

\documentclass[fleqn,usenatbib]{mnras}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	
\usepackage{amsmath}	
\usepackage{amssymb}
\usepackage{pdflscape}
\usepackage{siunitx}
\def\lya{Ly$\alpha$~}
\usepackage{color}
\definecolor{notecolor}{rgb}{0.8,0,0}
\newcommand{\gk}[1]{{\bf \color{notecolor} [#1]}}
\def\HI{\hbox{H~$\scriptstyle\rm I$}}
\def\HII{\hbox{H~$\scriptstyle\rm II$}}
\def\nHI{{\rm HI}}
\def\nH{{\rm H}}
\def\nHII{{\rm HII}}
\def\nHe{{\rm He}}
\def\nHeI{{\rm HeI}}
\def\nHeII{{\rm HeII}}
\def\nHeIII{{\rm HeIII}}
\def\bHII{\hbox{\bf H~$\scriptstyle\bf II$}}
\def\HeI{\hbox{He~$\scriptstyle\rm I$}}
\def\HeII{\hbox{He~$\scriptstyle\rm II$}}
\def\HeIII{\hbox{He~$\scriptstyle\rm III$}}
\def\bHeII{\hbox{\bf He~$\scriptstyle\bf II$}}
\def\HeIII{\hbox{He~$\scriptstyle\rm III$}}
\usepackage{dcolumn}
\newcolumntype{.}{D{.}{.}{2.7}}
\newcolumntype{e}{D{.}{.}{2.10}}
\newcolumntype{d}{D{.}{.}{2.3}}

@@

\author[Kulkarni et al.]
       {{Girish Kulkarni$^{1}$\thanks{Email: kulkarni@ast.cam.ac.uk},
           G\'abor Worseck$^{2}$
           and Joseph F.~Hennawi$^{3}$} \\
         $^1$Institute of Astronomy and Kavli Institute of Cosmology,
         University of Cambridge, Madingley Road, Cambridge CB3 0HA,
         UK \\
         $^2$Institute f\"ur Physik und Astronomie, Universit\"at
         Potsdam, Karl-Liebknecht-Stra\ss e\ 24/25, 14476 Potsdam,
         Germany \\
         $^3$Department of Physics, Broida Hall, UC Santa Barbara,
         Santa Barbara, CA 93106-9530 USA}
       \date{Accepted ---. Received ---; in original form ---}

\pubyear{2016}

@@

The luminosity function of active galactic nuclei (AGN) and its
evolution over cosmological time scales has been a matter of central
interest of a large body of work over the last four decades
\citep{1978A&A....68...17M, 1983ApJ...269..352S, 1988ApJ...325...92K,
  1988MNRAS.235..935B, 1993ApJ...406L..43H, 1994ApJ...421..412W,
  1995AJ....110...68S, 1995AJ....110.2553K, 1995ApJ...438..623P,
  2000MNRAS.317.1014B, 2001AJ....121...54F, 2004AJ....128..515F,
  2006AJ....131.2766R, 2007ApJ...654..731H, 2009MNRAS.392...19C,
  2010AJ....139..906W, 2011ApJ...728L..26G, 2013ApJ...773...14R,
  2013ApJ...768..105M, 2015AA...578A..83G, 2015ApJ...798...28K,
  2016ApJ...829...33Y, 2016ApJ...833..222J}.  Determination of the AGN
luminosity function constrains models of the build-up of supermassive
black holes \citep{1982MNRAS.200..115S, 2004ApJ...602..603Y,
  2007ApJ...654..731H, 2008ApJ...679..118S, 2009A&A...493...55E,
  2000ApJ...531...42H, 2010MNRAS.401.2531A, 2013ApJ...773...14R,
  2014ApJ...787...73D, 2015MNRAS.448.3603D, 2015MNRAS.452..575S,
  2016MNRAS.462..190R}.  Due to the incidence of supermassive black
holes in most galaxies, the tight scaling relations observed to exist
between the mass of these black holes and properties of their host
galaxies, and the increasing consensus that AGN activity feeds back on
the host galaxy evolution, the AGN luminosity function also constrains
models of galaxy formation \citep{2006ApJ...650...42L,
  2008MNRAS.385.1846M}.  Finally, thanks to their relative brightness
and high Lyman-continuum (LyC) escape fractions, the luminosity
function of AGN determines their contribution to the radiation
background that influences the temperature and ionization state of the
intergalactic medium (IGM), possibly even primarily driving hydrogen
and helium reionization in the last twelve billion years
\citep{2012ApJ...746..125H, 2015AA...578A..83G, 2017ApJ...847L..15O,
  2018MNRAS.474.2904P, 2018arXiv180104931P}.

There is a renewed interest in understanding the evolution of the UV
luminosity function of AGN, triggered by the discovery of 19
low-luminosity ($-18.9>M_{1450}>-22.6$) AGN candidates between
redshifts $z=4.1$ and $6.3$ by \citet{2015AA...578A..83G} using a
novel X-ray/NIR selection criterion.  This finding suggested that if
100\% LyC escape fraction is assumed AGN brighter than $M_{1450}=-18$
can potentially produce all of the metagalactic hydrogen
photoionization rate inferred from the \lya forest at $4<z<6$, despite
the fact that these AGN are fainter than the brightest star-forming
galaxies at these redshifts.  (The UV luminosity function of galaxies
at $z=6$ has $M^*_{1450}\sim -20$.)  A significant presence of AGN at
high redshift ($z\sim 6$) and a dominant contribution of AGN to
reionization is appealing as the LyC escape fraction of galaxies is
uncertain.  High-redshift galaxies down to rest-frame UV magnitude
$M_\mathrm{UV}=-12.5$ ($L\sim 10^{-3}L^*$) at $z=6$
\citep{2017ApJ...835..113L} and redshifts up to $z=11.1$
\citep{2016ApJ...819..129O} have now been reported.  But the escape of
LyC photons has been measured in only a few comparatively bright
($L>0.5L^*$) galaxies at relatively low redshifts ($z < 4$).  The
escape fraction in these galaxies reveals a broad distribution from
less than 2\% to more than 80\% \citep{2010ApJ...725.1011V,
  2011ApJ...736...41B, 2015ApJ...804...17S, 2015ApJ...810..107M,
  2016A&A...585A..48G, 2017MNRAS.468..389J, 2018arXiv180506071S,
  2018arXiv180601741F, 2018arXiv180511621M, 2018arXiv180303655C}, but
the average escape fraction is typically lower than 20\%.  Statistical
constraints from \HI\ column density measurements in gamma-ray burst
(GRB) afterglow spectra suggest an even lower escape fraction of 0.5\%
\citep{2007ApJ...667L.125C, 2009ApJS..185..526F, 2018arXiv180507318T}.
This is a challenge for reionization models, which require an escape
fraction of about 20\% in galaxies as faint as $M_\mathrm{UV}=-13$
\citep{2016PASA...33...37F, 2015ApJ...802L..19R, 2016MNRAS.457.4051K}.
An enhanced incidence of high-redshift AGN may also be consistent with
the shallow bright-end slopes of the $z\sim 7$ UV luminosity function
of galaxies relative to a Schechter function
\citep{2012MNRAS.426.2772B, 2014MNRAS.440.2810B, 2014ApJ...792...76B,
  2015MNRAS.452.1817B} and the hard spectra of these galaxies
\citep{2015MNRAS.450.1846S, 2015MNRAS.454.1393S, 2017MNRAS.464..469S}.
Finally, AGN may also provide a natural explanation for the large
scatter in the \lya opacity between different quasar sightlines close
to redshift $z=6$ \citep{2015MNRAS.447.3402B, 2018arXiv180208177B}.
%% JFH Please cite new paper by eilers et al. here, not yet on archive. 
These opacity fluctuations extend to substantially larger scales
($\gtrsim 50\, h^{-1}$cMpc) than expected in galaxies-dominated
reionization models (\citealt{2015MNRAS.453.2943C}; although see
\citealt{2016MNRAS.460.1328D, 2015ApJ...813L..38D,
  2018arXiv180308932B}).  AGN clustering can result in these
fluctuations naturally if there is a significant contribution
($\gtrsim 50\%$) of AGN to the ionising emissivity at $z=5$--$6$
\citep{2017MNRAS.465.3429C}.

Several counter-arguments against a higher incidence of AGN at high
redshift have also been made. \citet{2017MNRAS.468.4691D} and
\citet{2018MNRAS.473.1416M} pointed out that an enhanced number
density of AGN at $z\sim 6$ can lead to a much earlier
\HeII\ reionization.  For instance, He~\textsc{ii} reionization is
complete at $z=4.5$ in the model of \citet{2015ApJ...813L...8M},
compared to $z=3$ in the standard scenario
\citep{2012ApJ...746..125H}.  Such early \HeII\ reionization will also
result in higher IGM temperatures due to the associated photoheating.
\citet{2017MNRAS.468.4691D} found that the temperature of the IGM at
mean density is twice as much in AGN-dominated reionization models as
the standard models at $z=3.5$--$5$, in conflict with constraints from
the \lya forest.  This inconsistency could be avoided by postulating a
reduced escape fraction of \HeII-ionizing photons in AGN, but it is
difficult to reconcile this with a unit escape fraction of
hydrogen-ionizing photons that is required to explain the Ly$\alpha$
opacity fluctuations.  Further arguments against AGN-dominated
reionization have been presented by \citet{2016MNRAS.459.2299F}, who
analysed the simulations of metal-line absorbers at $z\sim 6$.
\citet{2016MNRAS.459.2299F} find that in their cosmological radiation
hydrodynamical simulations AGN-dominated UV background results in too
many C~\textsc{iv} absorption systems relative to Si~\textsc{iv} and
C~\textsc{ii} at $z\sim 6$.  (However, note that these simulations
assume a $L_\nu\propto\nu^{-1.57}$ AGN SED at extreme UV
\citep{2002ApJ...565..773T}, which somewhat harder than recent
measurements ($L_\nu\propto\nu^{-1.7}$) by
\citet{2015MNRAS.449.4204L}.  \citet{2016MNRAS.459.2299F} found that,
relative to the the standard model \citep{2012ApJ...746..125H}, the
observed N(Si~\textsc{iv})/N(C~\textsc{iv}) column density ratio
measurements seem to require a harder and more intense $>4$~Ry
background.)  Finally, comparing the \citet{2015AA...578A..83G} sample
to X-ray-selected quasar data at $z=0$--$6$,
\citet{2017MNRAS.465.1915R} argued that the faint end of the AGN UV
luminosity function at $z\sim 6$ is probably shallower that that
reported by \citet{2015AA...578A..83G}.  \citet{2017MNRAS.465.1915R}
argue that the apparent contradiction with the results of
\citet{2015AA...578A..83G} could be explained by contamination from
the host galaxies for faint AGN.

%---

@@

%% -- some people use photo-zs
%% -- some people mix binned data
%% -- some people fix slopes
%% -- some people exclude some data arbitrarily

@@

We included surveys based on a set of simple criteria:
\begin{enumerate}
\item High spectroscopic completeness of the target sample.
\item Accurate rest-frame UV-optical CCD photometry.
\item Statistical power (sample size, coverage in $z$ and/or absolute magnitude).
\item A well-characterised selection function.
\end{enumerate}
As a prerequisite for a joint analysis of the QLF we obtained the
survey selection functions in electronic form, either from the
publication or by request from the authors. As a reference for future surveys we
make them electronically available here
%% JFH rather than publish them here say ``we make them electronically available here''. Cite an appendix where
%% you do this. 
in modified and homogenised form
(Section~\ref{sect:datahom}).

@@

%% JFH This section is long tedious and hard to follow. Perhaps we
%% should put a brief description here, relegate these details to the
%% appendix, and in the appendix, structure this section by
%% sub-sections for each individual survey.

@@

%% JFH We need a better roadmap here for where this section is going. 1) Binned estimates 2) individual
%% fits in redshift bins. 3) examining redshift evolution. 4) attempting a joint fit.
%% GW: Agreed. Modified the roadmap.

@@

In a magnitude bin $[M_\mathrm{min}, M_\mathrm{max})$, and redshift
  bin $[z_\mathrm{min}, z_\mathrm{max})$, we define the binned luminosity
    %% JFH Notation error here. 
function as \citep{2000MNRAS.311..433P}
\begin{equation}
  \phi \equiv \frac{N_\mathrm{QSO}}{V_\mathrm{bin}},
\end{equation}
where $N_\mathrm{QSO}$ is the number of quasars with magnitude
$M_\mathrm{min}\leq M<M_\mathrm{max}$ and redshift
$z_\mathrm{min}\leq z<z_\mathrm{max}$, and
\begin{equation}
  V_\mathrm{bin} = \int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z},
  \label{eqn:vi}
\end{equation}
is the effective volume of the bin. The inclusion of the survey selection function
$f(M,z)$ (Section~\ref{sect:datahom}) in Equation~(\ref{eqn:vi}) accounts for
what are sometimes called ``incomplete bins''
\citep{2006AJ....131.2766R}.  The comoving volume element $\mathrm{d}V/\mathrm{d}z$ is
given by
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z}=\frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}\times A\times\frac{4\pi}{41253},
\end{equation}
where $A$ is the survey area in deg$^2$, and 
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}=\frac{c}{H_0}\frac{d_L^2\left(z\right)}
       {\left(1+z\right)^2\sqrt{\Omega_\mathrm{m}\left(1+z\right)^3+\Omega_\Lambda}}
  \label{eqn:dvdzdo}
\end{equation}
denotes the comoving volume element per unit solid angle
\citep{1999astro.ph..5116H}. The resulting luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.

@@

In each redshift bin, we model the quasar
luminosity function as a double power law 
\citep[e.g.][]{1988MNRAS.235..935B}
\begin{equation}
  \phi(M) =
  \frac{\phi_*}{10^{0.4(\alpha+1)(M-M_*)}+10^{0.4(\beta+1)(M-M_*)}}
  \label{eqn:dpl}
\end{equation}
with four free parameters: (i) the amplitude $\phi_*$, (ii) the break magnitude $M_*$,
(iii)  the bright-end slope $\alpha$, and (iv) the faint-end slope $\beta$.
%%GW: Maybe give the range of the uniform priors.
%% JFH SOmewhere indicate taht this procedure of ubninned LF estimates is standard practice. 
%% GW: Agreed. Maybe add references to Marshall et al. and other seminal papers further up in the text?
By assuming broad, uniform priors, we obtain posterior probability
distributions for these parameters using MCMC \citep[e.g.,][]{jaynes}.
The joint posterior probability distribution of the model parameters
is then written as
\begin{multline}
  p(\phi_*, M_*, \alpha, \beta | \{M_i, z_i\}) \propto \\ p(\phi_*, M_*,
  \alpha, \beta)p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta),
\end{multline}
where the constant of proportionality is independent of the luminosity
function parameters, and $\{M_i, z_i\}$ denotes the magnitudes and
redshifts of quasars falling in a redshift bin $[z_\mathrm{min},
  z_\mathrm{max})$.  We use a uniform prior distribution $p(\phi_*,
  M_*, \alpha, \beta)$ and assume that the likelihood
\begin{equation}
  \mathcal{L}\equiv p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta)
\end{equation}
is given by $\phi(M)$
%% JFH  ``is given by $\phi(M)$ itself''
with suitable normalisation.  The negative
logarithm of the likelihood $S\equiv -2\ln\mathcal{L}$ can then be
written as
\begin{multline}
  S = -2\sum_{i=1}^N\ln\phi(M_i, z_i)\\+2\int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, \phi(M,z) f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z}.
  \label{eqn:S}
\end{multline}
where $N$ is the total number of quasars in the redshift bin and the
luminosity integral in the second term on the right hand side is on
the surveyed range of $M$.  We use the \texttt{emcee} code
\citep{2013PASP..125..306F} for MCMC.

%% JFH SOmewhere indicate taht this procedure of ubninned LF estimates is standard practice. 
The above likelihood can also be understood as the limit of the
Poisson likelihood in luminosity and redshift bins
\citep{1983ApJ...269...35M, 2001AJ....121...54F}.  We can write the
probability of observing $n_{ij}$ quasars in the $(M_i, z_j)$ bin as
the Poisson distribution
\begin{equation}
  \mathcal{L}=\prod_{i,j}\frac{e^{-\mu_{ij}}\mu_{ij}^{n_{ij}}}{n_{ij}!},
  \label{eqn:lhood}
\end{equation}
where 
\begin{equation}
  \mu_{ij}= \int_{M_i}^{M_{i+1}}\mathrm{d}M\int_{z_j}^{z_{j+1}}\mathrm{d}z\, \phi(M,z) f(M, z)
  \,\frac{\mathrm{d}V}{\mathrm{d}z},
\end{equation}
is the average number of quasars expected in the $(M_i, z_j)$ bin
given the luminosity function $\phi(M,z)$.  In the limit of
infinitesimal bins, $n_{ij}=0$ or $1$, and Equation~(\ref{eqn:lhood})
can be simplified to obtain Equation~(\ref{eqn:S}).

>>>>

\documentclass[fleqn,usenatbib]{mnras}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{graphicx}	
\usepackage{amsmath}	
\usepackage{amssymb}
\usepackage{pdflscape}
\usepackage{siunitx}
\usepackage{color}
\definecolor{notecolor}{rgb}{0.8,0,0}
\newcommand{\gk}[1]{{\bf \color{notecolor} [#1]}}
\def\lya{Ly$\alpha$~}
\def\HI{\hbox{H~$\scriptstyle\rm I$}}
\def\HII{\hbox{H~$\scriptstyle\rm II$}}
\def\nHI{{\rm HI}}
\def\nH{{\rm H}}
\def\nHII{{\rm HII}}
\def\nHe{{\rm He}}
\def\nHeI{{\rm HeI}}
\def\nHeII{{\rm HeII}}
\def\nHeIII{{\rm HeIII}}
\def\bHII{\hbox{\bf H~$\scriptstyle\bf II$}}
\def\HeI{\hbox{He~$\scriptstyle\rm I$}}
\def\HeII{\hbox{He~$\scriptstyle\rm II$}}
\def\HeIII{\hbox{He~$\scriptstyle\rm III$}}
\def\bHeII{\hbox{\bf He~$\scriptstyle\bf II$}}
\def\HeIII{\hbox{He~$\scriptstyle\rm III$}}
\usepackage{dcolumn}
\newcolumntype{.}{D{.}{.}{2.7}}
\newcolumntype{e}{D{.}{.}{2.10}}
\newcolumntype{d}{D{.}{.}{2.3}}

@@

\author[Kulkarni et al.]
       {{Girish Kulkarni$^{1,2}$\thanks{Email: kulkarni@ast.cam.ac.uk},
           G\'abor Worseck$^{3}$
           and Joseph F.~Hennawi$^{4}$} \\
         $^1$Institute of Astronomy,
         University of Cambridge, Madingley Road, Cambridge CB3 0HA,
         UK \\
         $^2$Kavli Institute of Cosmology,
         University of Cambridge, Madingley Road, Cambridge CB3 0HA,
         UK \\
         $^2$Institute f\"ur Physik und Astronomie, Universit\"at
         Potsdam, Karl-Liebknecht-Stra\ss e\ 24/25, 14476 Potsdam,
         Germany \\
         $^3$Department of Physics, Broida Hall, UC Santa Barbara,
         Santa Barbara, CA 93106-9530, USA}
       \date{Accepted ---. Received ---; in original form ---}

\pubyear{2018}

@@

The luminosity function of active galactic nuclei (AGN) and its
evolution over cosmological time scales has been a matter of central
interest of a large body of work over the last four decades
\citep{1968ApJ...151..393S, 1978A&A....68...17M, 1983ApJ...269..352S,
  1988ApJ...325...92K, 1988MNRAS.235..935B, 1993ApJ...406L..43H,
  1994ApJ...421..412W, 1995AJ....110...68S, 1995AJ....110.2553K,
  1995ApJ...438..623P, 2000MNRAS.317.1014B, 2001AJ....121...54F,
  2004AJ....128..515F, 2006AJ....131.2766R, 2007ApJ...654..731H,
  2009MNRAS.392...19C, 2010AJ....139..906W, 2011ApJ...728L..26G,
  2013ApJ...773...14R, 2013ApJ...768..105M, 2015AA...578A..83G,
  2015ApJ...798...28K, 2016ApJ...829...33Y, 2016ApJ...833..222J}.
Determination of the AGN luminosity function constrains models of the
build-up of supermassive black holes \citep{1982MNRAS.200..115S,
  2004ApJ...602..603Y, 2007ApJ...654..731H, 2008ApJ...679..118S,
  2009A&A...493...55E, 2000ApJ...531...42H, 2010MNRAS.401.2531A,
  2013ApJ...773...14R, 2014ApJ...787...73D, 2015MNRAS.448.3603D,
  2015MNRAS.452..575S, 2016MNRAS.462..190R}.  Due to the incidence of
supermassive black holes in most galaxies, the tight scaling relations
observed to exist between the mass of these black holes and properties
of their host galaxies, and the increasing consensus that AGN activity
feeds back on the host galaxy evolution, the AGN luminosity function
also constrains models of galaxy formation \citep{2006ApJ...650...42L,
  2008MNRAS.385.1846M}.  Finally, thanks to their relative brightness
and high Lyman-continuum (LyC) escape fractions, the luminosity
function of AGN determines their contribution to the radiation
background that influences the temperature and ionization state of the
intergalactic medium (IGM), possibly even primarily driving hydrogen
and helium reionization in the last twelve billion years
\citep{2012ApJ...746..125H, 2015AA...578A..83G, 2017ApJ...847L..15O,
  2018MNRAS.474.2904P, 2018arXiv180104931P}.

There is a renewed interest in understanding the evolution of the UV
luminosity function of AGN, triggered by the discovery of 19
low-luminosity ($-18.9>M_{1450}>-22.6$) AGN candidates between
redshifts $z=4.1$ and $6.3$ by \citet{2015AA...578A..83G} using a
novel X-ray/NIR selection criterion.  This finding suggested that if
100\% LyC escape fraction is assumed AGN brighter than $M_{1450}=-18$
can potentially produce all of the metagalactic hydrogen
photoionization rate inferred from the \lya forest at $4<z<6$, despite
the fact that these AGN are fainter than the brightest star-forming
galaxies at these redshifts.  (The UV luminosity function of galaxies
at $z=6$ has $M^*_{1450}\sim -20$.)  A significant presence of AGN at
high redshift ($z\sim 6$) and a dominant contribution of AGN to
reionization is appealing as the LyC escape fraction of galaxies is
uncertain.  High-redshift galaxies down to rest-frame UV magnitude
$M_\mathrm{UV}=-12.5$ ($L\sim 10^{-3}L^*_\mathrm{galaxies}$) at $z=6$
\citep{2017ApJ...835..113L} and redshifts up to $z=11.1$
\citep{2016ApJ...819..129O} have now been reported.  But the escape of
LyC photons has been measured in only a few comparatively bright
($L>0.5L^*_\mathrm{galaxies}$) galaxies at relatively low redshifts
($z < 4$).  The escape fraction in these galaxies reveals a broad
distribution from less than 2\% to more than 80\%
\citep{2010ApJ...725.1011V, 2011ApJ...736...41B, 2015ApJ...804...17S,
  2015ApJ...810..107M, 2016A&A...585A..48G, 2017MNRAS.468..389J,
  2018arXiv180506071S, 2018arXiv180601741F, 2018arXiv180511621M,
  2018arXiv180303655C}, but the average escape fraction is typically
lower than 20\%.  Statistical constraints from \HI\ column density
measurements in gamma-ray burst (GRB) afterglow spectra suggest an
even lower escape fraction of 0.5\% \citep{2007ApJ...667L.125C,
  2009ApJS..185..526F, 2018arXiv180507318T}.  This is a challenge for
reionization models, which require an escape fraction of about 20\% in
galaxies as faint as $M_\mathrm{UV}=-13$ \citep{2016PASA...33...37F,
  2015ApJ...802L..19R, 2016MNRAS.457.4051K}.  An enhanced incidence of
high-redshift AGN may also be consistent with the shallow bright-end
slopes of the $z\sim 7$ UV luminosity function of galaxies relative to
a Schechter function \citep{2012MNRAS.426.2772B, 2014MNRAS.440.2810B,
  2014ApJ...792...76B, 2015MNRAS.452.1817B} and the hard spectra of
these galaxies \citep{2015MNRAS.450.1846S, 2015MNRAS.454.1393S,
  2017MNRAS.464..469S}.  Finally, AGN may also provide a natural
explanation for the large scatter in the \lya opacity between
different quasar sightlines close to redshift $z=6$
\citep{2015MNRAS.447.3402B, 2018arXiv180208177B}.
These opacity fluctuations extend to substantially larger scales
($\gtrsim 50\, h^{-1}$cMpc) than expected in galaxies-dominated
reionization models (\citealt{2015MNRAS.453.2943C}; although see
\citealt{2016MNRAS.460.1328D, 2015ApJ...813L..38D,
  2018arXiv180308932B}).  AGN clustering can result in these
fluctuations naturally if there is a significant contribution
($\gtrsim 50\%$) of AGN to the ionising emissivity at $z=5$--$6$
\citep{2017MNRAS.465.3429C}.

Several counter-arguments against a higher incidence of AGN at high
redshift have also been made. \citet{2017MNRAS.468.4691D} and
\citet{2018MNRAS.473.1416M} pointed out that an enhanced number
density of AGN at $z\sim 6$ can lead to a much earlier
\HeII\ reionization (see also \citealt{2017MNRAS.471..255K}).  For
instance, He~\textsc{ii} reionization is complete at $z=4.5$ in the
model of \citet{2015ApJ...813L...8M}, compared to $z=3$ in the
standard scenario \citep{2012ApJ...746..125H}.  Such early
\HeII\ reionization will also result in higher IGM temperatures due to
the associated photoheating.  \citet{2017MNRAS.468.4691D} found that
the temperature of the IGM at mean density is twice as much in
AGN-dominated reionization models as the standard models at
$z=3.5$--$5$, in conflict with constraints from the \lya forest.  This
inconsistency could be avoided by postulating a reduced escape
fraction of \HeII-ionizing photons in AGN, but it is difficult to
reconcile this with a unit escape fraction of hydrogen-ionizing
photons that is required to explain the Ly$\alpha$ opacity
fluctuations.  Further arguments against AGN-dominated reionization
have been presented by \citet{2016MNRAS.459.2299F}, who analysed the
simulations of metal-line absorbers at $z\sim 6$.
\citet{2016MNRAS.459.2299F} find that in their cosmological radiation
hydrodynamical simulations AGN-dominated UV background results in too
many C~\textsc{iv} absorption systems relative to Si~\textsc{iv} and
C~\textsc{ii} at $z\sim 6$.  (However, note that these simulations
assume a $L_\nu\propto\nu^{-1.57}$ AGN SED at extreme UV
\citep{2002ApJ...565..773T}, which is somewhat harder than recent
measurements ($L_\nu\propto\nu^{-1.7}$) by
\citet{2015MNRAS.449.4204L}.  \citet{2016MNRAS.459.2299F} found that,
relative to the the standard model \citep{2012ApJ...746..125H}, the
observed N(Si~\textsc{iv})/N(C~\textsc{iv}) column density ratio
measurements seem to require a harder and more intense $>4$~Ry
background.)  Finally, comparing the \citet{2015AA...578A..83G} sample
to X-ray-selected quasar data at $z=0$--$6$,
\citet{2017MNRAS.465.1915R} argued that the faint end of the AGN UV
luminosity function at $z\sim 6$ is probably shallower that that
reported by \citet{2015AA...578A..83G}.  \citet{2017MNRAS.465.1915R}
argue that the apparent contradiction with the results of
\citet{2015AA...578A..83G} could be explained by contamination from
the host galaxies for faint AGN \citep[see
  also][]{2015MNRAS.453.1946G,2015MNRAS.448.3167W,2016MNRAS.463..348V}.

@@

@@

We included surveys based on a set of simple criteria:
\begin{enumerate}
\item High spectroscopic completeness of the target sample.
\item Accurate rest-frame UV-optical CCD photometry.
\item Statistical power (sample size, coverage in $z$ and/or absolute magnitude).
\item A well-characterised selection function.
\end{enumerate}
As a prerequisite for a joint analysis of the QLF we obtained the
survey selection functions in electronic form, either from the
publication or by request from the authors. As a reference for future
surveys we make them electronically available here in modified and
homogenised form (see Section~\ref{sect:datahom} and
Appendix~\ref{sec:code}).

@@

@@

@@

In a magnitude bin $[M_\mathrm{min}, M_\mathrm{max})$, and redshift
  bin $[z_\mathrm{min}, z_\mathrm{max})$, we define the binned luminosity
    %% JFH Notation error here. 
function as \citep{2000MNRAS.311..433P}
\begin{equation}
  \phi \equiv \frac{N_\mathrm{QSO}}{V_\mathrm{bin}},
\end{equation}
where $N_\mathrm{QSO}$ is the number of quasars with magnitude
$M_\mathrm{min}\leq M<M_\mathrm{max}$ and redshift
$z_\mathrm{min}\leq z<z_\mathrm{max}$, and
\begin{equation}
  V_\mathrm{bin} = \int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z},
  \label{eqn:vi}
\end{equation}
is the effective volume of the bin. The inclusion of the survey selection function
$f(M,z)$ (Section~\ref{sect:datahom}) in Equation~(\ref{eqn:vi}) accounts for
what are sometimes called ``incomplete bins''
\citep{2006AJ....131.2766R}.  The comoving volume element $\mathrm{d}V/\mathrm{d}z$ is
given by
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z}=\frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}\times A\times\frac{4\pi}{41253},
\end{equation}
where $A$ is the survey area in deg$^2$, and 
\begin{equation}
  \frac{\mathrm{d}V}{\mathrm{d}z\,\mathrm{d}\Omega}=\frac{c}{H_0}\frac{d_L^2\left(z\right)}
       {\left(1+z\right)^2\left[\Omega_\mathrm{m}\left(1+z\right)^3+\Omega_\Lambda\right]^{1/2}}
  \label{eqn:dvdzdo}
\end{equation}
denotes the comoving volume element per unit solid angle
\citep{1999astro.ph..5116H}. The resulting luminosity function $\phi$ has units of
$\mathrm{cMpc}^{-3}\mathrm{mag}^{-1}$.

@@

In each redshift bin, we model the quasar
luminosity function as a double power law 
\citep[e.g.][]{1988MNRAS.235..935B}
\begin{equation}
  \phi(M) =
  \frac{\phi_*}{10^{0.4(\alpha+1)(M-M_*)}+10^{0.4(\beta+1)(M-M_*)}}
  \label{eqn:dpl}
\end{equation}
with four free parameters: (i) the amplitude $\phi_*$, (ii) the break magnitude $M_*$,
(iii)  the bright-end slope $\alpha$, and (iv) the faint-end slope $\beta$.
%%GW: Maybe give the range of the uniform priors.
%% JFH SOmewhere indicate taht this procedure of ubninned LF estimates is standard practice. 
%% GW: Agreed. Maybe add references to Marshall et al. and other seminal papers further up in the text?
By assuming broad, uniform priors, we obtain posterior probability
distributions for these parameters using the Markov Chain Monte Carlo technique \citep[MCMC, e.g.,][]{jaynes}.
The joint posterior probability distribution of the model parameters
is then written as
\begin{multline}
  p(\phi_*, M_*, \alpha, \beta | \{M_i, z_i\}) \propto \\ p(\phi_*, M_*,
  \alpha, \beta)p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta),
\end{multline}
where the constant of proportionality is independent of the luminosity
function parameters, and $\{M_i, z_i\}$ denotes the magnitudes and
redshifts of quasars falling in a redshift bin $[z_\mathrm{min},
  z_\mathrm{max})$.  We use a uniform prior distribution $p(\phi_*,
  M_*, \alpha, \beta)$ and assume that the likelihood
\begin{equation}
  \mathcal{L}\equiv p(\{M_i, z_i\} | \phi_*, M_*, \alpha, \beta)
\end{equation}
is given by $\phi(M)$
%% JFH  ``is given by $\phi(M)$ itself''
with suitable normalisation.  The negative
logarithm of the likelihood $S\equiv -2\ln\mathcal{L}$ can then be
written as
\begin{multline}
  S = -2\sum_{i=1}^{N_\mathrm{QSO}}\ln\phi(M_i, z_i)\\+2\int_{M_\mathrm{min}}^{M_\mathrm{max}}\mathrm{d}M
  \int_{z_\mathrm{min}}^{z_\mathrm{max}}\mathrm{d}z\, \phi(M,z) f(M, z)\,\frac{\mathrm{d}V}{\mathrm{d}z},
  \label{eqn:S}
\end{multline}
where the integral over magnitude is on the surveyed range of $M$.  We
use the \texttt{emcee} code \citep{2013PASP..125..306F} for MCMC.

The above likelihood can also be understood as the limit of the
Poisson likelihood in luminosity and redshift bins
\citep{1983ApJ...269...35M, 2001AJ....121...54F}.  We can write the
probability of observing $n_{ij}$ quasars in the $(M_i, z_j)$ bin as
the Poisson distribution
\begin{equation}
  \mathcal{L}=\prod_{i,j}\frac{e^{-\mu_{ij}}\mu_{ij}^{n_{ij}}}{n_{ij}!},
  \label{eqn:lhood}
\end{equation}
where 
\begin{equation}
  \mu_{ij}= \int_{M_i}^{M_{i+1}}\mathrm{d}M\int_{z_j}^{z_{j+1}}\mathrm{d}z\, \phi(M,z) f(M, z)
  \,\frac{\mathrm{d}V}{\mathrm{d}z},
\end{equation}
is the average number of quasars expected in the $(M_i, z_j)$ bin
given the luminosity function $\phi(M,z)$.  In the limit of
infinitesimal bins, $n_{ij}=0$ or $1$, and Equation~(\ref{eqn:lhood})
can be simplified to obtain Equation~(\ref{eqn:S}).
